\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{geometry}
\usepackage{tcolorbox}
\usepackage{fancyhdr}
\usepackage{hyperref}
\geometry{margin=1in}

% Custom colors
\definecolor{statecol}{RGB}{70,130,180}
\definecolor{transitioncol}{RGB}{0,100,0}

\title{\Large\textbf{CSL253 - Theory of Computation}\\[0.5em]
	\large Tutorial 7}
\author{}
\date{}


\newtheorem{theorem}{Theorem}

\begin{document}
	
	\maketitle
	
	% Team members in a nice box
	\begin{tcolorbox}[
		title=\textbf{Team Members},
		colback=white,
		colframe=statecol,
		arc=0mm
		]
		\begin{enumerate}[leftmargin=*]
			\item Kartikeya Nainakhwal - 12341090
			\item Paritosh Lahre - 12341550
			\item Rahul Dev Reddy - 12342390
		\end{enumerate}
	\end{tcolorbox}
	
	% Question in a colored box
	\begin{tcolorbox}[
		title=\textbf{Question: 2},
		colback=white,
		colframe=transitioncol,
		arc=0mm
		]
		Let 
		\[
		A_{ECFG} = \{\langle G \rangle \mid G \text{ is a CFG and } \varepsilon \in L(G)\}.
		\]
		Show that $A_{ECFG}$ is decidable.
	\end{tcolorbox}
	
	\vspace{1em}
	
	% Solution Sections
	\section*{Problem Statement}
	We define the language
	\[
	A_{ECFG} = \{\langle G\rangle \mid G \text{ is a context-free grammar and } \varepsilon \in L(G)\}.
	\]
	Prove that $A_{ECFG}$ is decidable; that is, there exists a Turing machine that, given any encoding $\langle G\rangle$, halts and correctly decides whether $G$ generates the empty string $\varepsilon$.
	
	\section*{Proof Idea}
	The key observation is that a CFG $G$ generates $\varepsilon$ precisely when its start symbol $S$ is ``nullable,'' meaning there exists a derivation $S \Rightarrow^*_R \varepsilon$.  We can compute the set of all nullable nonterminals by a simple fixpoint algorithm:
	\begin{enumerate}[label=\arabic*.]
		\item Initialize \textsf{Null} with any nonterminal that has a production directly deriving $\varepsilon$.
		\item Repeatedly add any nonterminal $A$ if there is a production
		\[A \to X_1 X_2 \cdots X_k\]
		such that each $X_i$ is either $\varepsilon$ or already in \textsf{Null}.
		\item Stop when no new symbols can be added.  At termination, $S\in\textsf{Null}$ iff $\varepsilon\in L(G)$.
	\end{enumerate}
	
	\section*{Proof}
	*	$A_{ECFG}$ is decidable.
	\begin{proof}
		We construct a deterministic Turing machine $M$ that decides $A_{ECFG}$.  On input $\langle G\rangle$, where $G=(V,\Sigma,R,S)$:
		\begin{enumerate}[label=\arabic*.]
			\item Parse the input into $(V,\Sigma,R,S)$; if malformed, \textbf{reject}.
			\item Let
			\[
			\mathrm{Null} \leftarrow \{A\in V \mid A\to\varepsilon\in R\}.
			\]
			\item \textbf{Repeat} the following until fixpoint:
			\begin{enumerate}[label=\alph*.]
				\item For each production $A\to X_1X_2\cdots X_k$ in $R$, if $\forall i,\;X_i=\varepsilon$ or $X_i\in\mathrm{Null}$, add $A$ to $\mathrm{Null}$.
			\end{enumerate}
			\item If $S\in\mathrm{Null}$, \textbf{accept}; otherwise, \textbf{reject}.
		\end{enumerate}
		
		\textbf{Correctness:}
		\begin{itemize}
			\item \emph{Soundness}: If $M$ accepts, then $S\in\mathrm{Null}$, so by construction there is a derivation $S\Rightarrow^*_R\varepsilon$, i.e.\ $\varepsilon\in L(G)$.
			\item \emph{Completeness}: If $\varepsilon\in L(G)$, any leftmost derivation shows each nonterminal on the path to $\varepsilon$ must be nullable.  By induction, our loop adds each such nonterminal to $\mathrm{Null}$, including $S$, so $M$ accepts.
			\item \emph{Termination}: Each iteration can only add new symbols to the finite set $\mathrm{Null}$ and proceeds until no additions are possible.  Since $|V|$ bounds the additions, $M$ halts after at most $|V|+1$ iterations.  All other steps are finite.  Thus $M$ always halts.
		\end{itemize}
	\end{proof}
	
	\section*{Conclusion}
	We have exhibited a concrete algorithm that, given any CFG $G$, computes whether $\varepsilon$ is in its language by determining nullable nonterminals.  The procedure halts on every input and correctly decides membership in $A_{ECFG}$.  Hence, $A_{ECFG}$ is decidable.
	
\end{document}
